---
title: "Introduction to the BARplot function"
author: 
  - Ann-Sophie Buchardt
  - Karl Bang Christensen
  - Sidsel Normann Jensen
date: "`r format(Sys.time(), '%B %d, %Y')`"
output: 
  rmarkdown::html_vignette:
    toc: true
    toc_depth: 3
vignette: >
  %\VignetteIndexEntry{Introduction to the BARplot function}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = FALSE,
  comment = "#>",
  #echo = TRUE, message=FALSE,
  tidy = TRUE, 
  tidy.opts = list(blank = TRUE, arrow = TRUE), 
  highlight = TRUE,
  #collapse = FALSE,
  #cache.extra = R.version, autodep=TRUE,
  fig.width=6, fig.height=4, comment=NA,
  fig.align="center"
  )

OkabeIto <- c("#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7", "#999999")

library(ggplot2)

theme_set(theme_minimal())

options(ggplot2.continuous.colour = OkabeIto)
options(ggplot2.continuous.fill = OkabeIto)
options(ggplot2.discrete.colour = OkabeIto)
options(ggplot2.discrete.fill = OkabeIto)
```

## Introduction

This vignette describes how one can use the `BARplot` function to visualise the distribution of item responses.

First, we load the `RASCHplot` package

```{r setup}
library(RASCHplot)
```

The functions in the package are applicable on polytomous as well as dichotomous items. 
We load the polytomous SPADI data set from the package. Users can either load their own data or use those saved in the workspace.

```{r}
data(SPADI)
```

The command loads data from this saved R data archive. Data are from a Danish validation study containing responses to the shoulder pain and disability index (SPADI) from 228 patients with rotator-cuff related disorders. 

```{r}
str(SPADI)
```

Let us focus on the eight disability items. 

```{r}
itemsIncomp <- SPADI[,9:16]
```

We can visualise the distribution of responses using the most basic call to `BARplot()`.

```{r, results = 'hide'}
BARplot(itemsIncomp)
```

## Arguments

```{r, results = 'hide'}
BARplot(itemsIncomp)
```

```{r, fig.height=1.5}
BARplot(itemsIncomp, which.item = 1)
```

```{r, results = 'hide'}
BARplot(itemsIncomp, freq = FALSE)
```

```{r, results = 'hide'}
BARplot(itemsIncomp, na.action = "na.rm")
```

```{r, results = 'hide'}
BARplot(itemsIncomp, addsums = TRUE)
```

```{r}
strat.var.gender <- factor(SPADI$gender)
levels(strat.var.gender) <-  list("male" = 1, "female" = 2)
```

```{r, results = 'hide', fig.height=7.5}
BARplot(itemsIncomp, strat.var = strat.var.gender)
```

```{r, fig.height=2}
BARplot(itemsIncomp, which.item = 2, strat.var = strat.var.gender)
```

```{r, fig.height=2}
BARplot(itemsIncomp, which.item = 2, freq = FALSE, strat.var = strat.var.gender)
```

## Working with `ggplot2`

Default theme for `ggplot2`

```{r, results = 'hide'}
theme_set(theme_gray())
BARplot(itemsIncomp)
```

```{r}
theme_set(theme_minimal())
```

```{r, results = 'hide'}
barobj <- BARplot(itemsIncomp)
```

```{r, results = 'hide'}
barobj + 
  scale_fill_brewer(palette="Set3", na.value = "grey")
```

```{r, results = 'hide'}
barobj + theme(legend.position="top")
```

```{r, results = 'hide'}
barobj + ggtitle("Distribution of item responses")
```
