---
title: "Introduction to the Yen's Q3 functions"
author: 
  - Ann-Sophie Buchardt
  - Karl Bang Christensen
date: "`r format(Sys.time(), '%B %d, %Y')`"
output: 
  rmarkdown::html_vignette:
    toc: true
    toc_depth: 3
vignette: >
  %\VignetteIndexEntry{Introduction to the Yen's Q3 functions}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = FALSE,
  comment = "#>",
  #echo = TRUE, message=FALSE,
  tidy = TRUE, 
  tidy.opts = list(blank = TRUE, arrow = TRUE), 
  highlight = TRUE,
  #collapse = FALSE,
  #cache.extra = R.version, autodep=TRUE,
  fig.width=6, fig.height=5, comment=NA,
  fig.align="center"
  )

library(ggsci)
jama <- pal_jama(palette = c("default"), alpha = 1)(7)

library(ggplot2)

theme_set(theme_minimal())

options(ggplot2.continuous.colour = jama)
options(ggplot2.continuous.fill = jama)
options(ggplot2.discrete.colour = jama)
options(ggplot2.discrete.fill = jama)
```

## Introduction

This R tutorial describes how to compute and visualize Yen's $Q_{3}$ statistics using R software and the `RASCHplot` package. In particular, it illustrates the usage of the following functions:

- `Q3()`
- `ggQ3star()`

First, we load the `RASCHplot` package:

```{r setup}
library(RASCHplot)
```

The functions in the package are applicable on polytomous as well as dichotomous items. 
We load the polytomous SPADI data set from the package. Users can either load their own data or use those saved in the workspace.

## Prepare the data

The `SPADI` is used:

```{r}
data(SPADI)

head(SPADI)
```

We remove incomplete cases:

```{r}
SPADI.complete <- SPADI[complete.cases(SPADI), ]
```

We create an object containing only items:

```{r}
it.SPADI <- SPADI.complete[, 9:16]
```

## Compute Yen's Q3 statistics

Yen's Q3 statistics are computed using the `Q3()` function:

```{r}
q3obj <- Q3(items = it.SPADI, method.item = "CML", method.person = "WML", model = "RMP")
```

Here CML is used for estimating item parameters and WML is used for estimating person parameters. Other options are available, see `?Q3`.

The `q3obj` includes the following values:

```{r}
names(q3obj)
```

We can, e.g., extract the $Q_{3\ast$ value from the `q3obj` object:

```{r}
q3obj$Q3star
```

## Visualise Yen's Q3 statistics

The `ggQ3star()` function is used to visualise Yen's Q3 statistics:

```{r}
ggQ3star(object = q3obj)
```

Below, we give examples of how to change the appearance of the plot.

As we saw above, the $Q_{3\ast}$ value is obtained for items D4 and D5:

```{r}
q3obj$Q3star
```

We observe that $Q_{3\ast} > 0.2$ and with the reasoning that these items hold the same information, we create a new variable as the sum of D4 and D5:

```{r}
it.SPADI$D4D5 <- it.SPADI$D4 + it.SPADI$D5
```

and remove the two:

```{r}
it.SPADI.2 <- it.SPADI[,-c(4,5)]
```

We compute Yen's Q3 statistics for this new set of items:

```{r}
q3obj2 <- Q3(items = it.SPADI.2, method.item = "CML", method.person = "WML", model = "RMP")
```

Again, we visualise Yen's Q3 statistics using `ggQ3star()`:

```{r}
ggQ3star(object = q3obj2)
```

We see that $Q_{3\ast} < 0.2$.


## Appearance

```{r}
ggQ3star(object = q3obj, title = "Yen's Q3(*) statistics")
```

```{r}
ggQ3star(object = q3obj, colors = c("chartreuse", "gray", "orchid"))
```


```{r}
ggQ3star(object = q3obj, lab_col = "goldenrod")
```


```{r}
ggQ3star(object = q3obj, method = "square")
```

```{r}
ggQ3star(object = q3obj, type = "upper", labels = "lower")
```

```{r}
ggQ3star(object = q3obj, type = "full")
```


```{r}
ggQ3star(object = q3obj, type = "full", markQ3star = "pch")
```

## Appearance and `ggplot2`

```{r}
ggQ3star(object = q3obj) +
  theme_void()
```

```{r}
ggQ3star(object = q3obj) + 
  theme_dark() +
  theme(axis.title.x = element_blank(),
        axis.title.y = element_blank(), 
        panel.grid.minor = element_blank())
```


## Combination

```{r}
ggQ3star(object = q3obj, lab_col = "white", outline.color = "white", outline.width = .5) + 
  theme_dark() +
  theme(axis.title.x = element_blank(),
        axis.title.y = element_blank(), 
        panel.grid.minor = element_blank())
```

