---
title: "Introduction to the RASCHplot package"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to the RASCHplot package}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  #echo = TRUE, message=FALSE,
  tidy = TRUE, 
  tidy.opts = list(blank = TRUE, arrow = TRUE), 
  highlight = TRUE,
  #collapse = FALSE,
  #cache.extra = R.version, autodep=TRUE,
  fig.width=6, fig.height=4, comment=NA,
  fig.align="center"
  )

library(tidyverse)

OkabeIto <- c("#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7", "#999999")

theme_set(theme_minimal())

options(ggplot2.continuous.colour = OkabeIto)
options(ggplot2.continuous.fill = OkabeIto)
options(ggplot2.discrete.colour = OkabeIto)
options(ggplot2.discrete.fill = OkabeIto)
```

# Introduction

Visualisation of Rasch item fit using a Conditional Item Characteristic Curve (CICC), which is a curve describing the expected item
mean as a function of the total score. It is possible to make an empirical CICC based on
the observed data as both the empirical expected item score and the total scores can be
calculated from the data. This emperical curve can then be compared to the model-based
CICC to visualize item fit.

Here, you will learn how to use the `RASCHplot` package to visualise responses to tests or questionnaires as well as Rasch model fit.

# Installation

```{r, eval=FALSE}
devtools::install_github("ERRTG/RASCHplot")
```


# Quick Start

The aim of this section is to give users a general sense of the package, including the functions, what they do and basic usage. We briefly introduce the functions, see the basic operations and have a look at the outputs. Hopefully, this will give users an idea about the functions available and their purpose, and where to find more information. More details are given in later sections.


First, we load the `RASCHplot` package

```{r setup}
library(RASCHplot)
```

The functions in the package are applicable on polytomous as well as dichotomous items. 
We load the polytomous SPADI data set from the package. Users can either load their own data or use those saved in the workspace.

```{r}
data(SPADI)
```

The command loads data from this saved R data archive. Data are from a Danish validation study containing responses to the shoulder pain and disability index (SPADI) from 228 patients with rotator-cuff related disorders. 

```{r}
str(SPADI)
```

Let us focus on the eight disability items. 

```{r}
itemsIncomp <- SPADI[,9:16]
```

We can visualise the distribution of responses using the most basic call to `BARplot()`.

```{r}
BARplot(itemsIncomp)
```

This produces a bar plot where the length of the bar  is proportional to 

Visualising Rasch model fit is main motivation for the package. 

```{r}
SPADI.complete <- SPADI[complete.cases(SPADI), ]
items <- SPADI.complete[, 9:16]
```


We fit a partial credit model using the `PCM` function from the `eRm` package.

```{r}
fit <- eRm::PCM(items)
```

We use the most basic call to `CICCplot()` to visualise the model fit of the first item.

```{r, results = 'hide'}
CICCplot(fit)
```

One of the key benefits of the `CICCplot()` function is the ability to impose a grouping on
the total score. This is helpful in choosing the plot that best communicates the level of fit or
misfit.

```{r, results = 'hide'}
CICCplot(fit, lower.groups = c(0,7,13,14,19,22,27,30,35))
```


```{r}
strat.vars <- list(gender = SPADI.complete[, "gender"])
DIFplot(model = fit, strat.vars = strat.vars)
```

# Function: `BARplot()`

## Arguments

```{r, results = 'hide'}
BARplot(itemsIncomp)
```

```{r, fig.height=1.5}
BARplot(itemsIncomp, which.item = 1)
```

```{r, results = 'hide'}
BARplot(itemsIncomp, freq = FALSE)
```

```{r, results = 'hide'}
BARplot(itemsIncomp, na.action = "na.rm")
```

```{r, results = 'hide'}
BARplot(itemsIncomp, addsums = TRUE)
```

```{r}
strat.var.gender <- factor(SPADI$gender)
levels(strat.var.gender) <-  list("male" = 1, "female" = 2)
```

```{r, results = 'hide', fig.height=7.5}
BARplot(itemsIncomp, strat.var = strat.var.gender)
```

```{r, fig.height=2}
BARplot(itemsIncomp, which.item = 2, strat.var = strat.var.gender)
```

```{r, fig.height=2}
BARplot(itemsIncomp, which.item = 2, freq = FALSE, strat.var = strat.var.gender)
```

## Working with `ggplot2`

Default theme for `ggplot2`

```{r, results = 'hide'}
theme_set(theme_gray())
BARplot(itemsIncomp)
```

```{r}
theme_set(theme_minimal())
```

```{r, results = 'hide'}
barobj <- BARplot(itemsIncomp)
```

```{r, results = 'hide'}
barobj + 
  scale_fill_brewer(palette="Set3", na.value = "grey")
```

```{r, results = 'hide'}
barobj + theme(legend.position="top")
```

```{r, results = 'hide'}
barobj + ggtitle("Distribution of item responses")
```




# Function: `CICCplot()`

## Arguments

```{r, results = 'hide'}
CICCplot(fit)
```

```{r, results = 'hide'}
CICCplot(fit, lower.groups = c(1,5,15,20,30,35))
```

```{r, results = 'hide'}
CICCplot(fit, error.bar = FALSE)
```

```{r, results = 'hide'}
CICCplot(fit, which.item = c(3,4))
```

```{r, results = 'hide'}
CICCplot(fit, all.items = TRUE)
```

Construc CICC plot for all items and grid the items in two columns

```{r, results = 'hide', fig.height=16}
CICCplot(model = fit, all.items = TRUE, grid.items = TRUE,
          common.legend = TRUE, ncol = 2, nrow = 4)
```

Construct CICC plot for items 1 and 2

```{r, results = 'hide'}
 p <- CICCplot(model = fit, which.item = c(1,2))
 p
 # Change colours for item 1
 p[[1]] +
   scale_colour_manual(values = c("burlywood","cornflowerblue"))
 # Change colours for item 2
 p[[2]] +
   scale_colour_manual(values = c("burlywood","cornflowerblue"))
```


## Working with `ggplot2`

Default theme for `ggplot2`

```{r, results = 'hide'}
theme_set(theme_gray())
CICCplot(fit)
```

```{r}
theme_set(theme_minimal())
```

```{r, results = 'hide'}
ciccobj <- CICCplot(fit)[[1]]
```

```{r, results = 'hide'}
ciccobj +
   scale_colour_manual(values = c("burlywood","cornflowerblue"))
```

Change colours and x axis limits

```{r, results = 'hide'}
ciccobj + 
  scale_colour_manual(values = c("burlywood","cornflowerblue")) +
  xlim(c(3, 37))
```


Change y axis limits

```{r, results = 'hide'}
ciccobj + 
  ylim(c(0,5))
```

```{r}
ciccobj +
  scale_x_continuous(breaks = seq(0, 40, 5), minor_breaks = NULL) +
  scale_y_continuous(breaks = 0:5, minor_breaks = NULL)   + 
  ylim(c(0,5))
```

```{r, results = 'hide'}
ciccobj + 
  ylab("Item-score for item D1") + 
  ggtitle("")
```

```{r}
ciccobj + 
  theme(plot.title = element_text(hjust = 0.5))
```




# Function: `DIFplot()`

## Arguments

```{r}
DIFplot(fit)
```

```{r}
strat.vars <- list(gender = SPADI.complete[, "gender"])
DIFplot(model = fit, strat.vars = strat.vars)
```

```{r}
strat.vars <- list(gender = SPADI.complete[, "gender"], over60 = SPADI.complete[, "over60"])
DIFplot(model = fit, strat.vars = strat.vars)
```

```{r}
DIFplot(model = fit, which.item = c(1,2), strat.vars = strat.vars)
```

```{r}
lower.groups <- list(gender = c(0, 1, 2, 5, 8, 10), over60 = c(0, 3, 6, 9))
DIFplot(model = fit, strat.vars = strat.vars, lower.groups = lower.groups)
```

```{r}
DIFplot(model = fit, which.item = c(1,2),
        strat.vars = strat.vars, lower.groups = lower.groups)
```

```{r}
lower.groups <- list(gender = list("1" = c(0, 1, 2, 5, 8, 10), "2" = c(0,  5, 10)),
                     over60 = list("0" = c(0, 1, 2, 3, 5),
                                   "1" = c(0, 1, 2, 5, 7, 8, 9, 10)))
DIFplot(model = fit, strat.vars = strat.vars, lower.groups = lower.groups)
```

